# libraries

```{r}
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(car)
library(broom)
library(ggcorrplot)
library(caret)
library(randomForest)
library(ipred)
library(stringr)
library(forcats)
```

# reading

```{r}
train_data <- read.csv("/Users/ishujaswani/Downloads/Columbia_Courses/App Frameworks 1/Final project/Spotify_PAC/analysisData.csv")
test_data <- read.csv("/Users/ishujaswani/Downloads/Columbia_Courses/App Frameworks 1/Final project/Spotify_PAC/scoringData.csv")
```

# data transformation

```{r}
library(stringr)

# train data genre column splits to 1 and 0

train_data$genre<- gsub("\\[|\\]","",as.character(train_data$genre))
train_data$genre <- ifelse(train_data$genre == "","No_Genre", train_data$genre)
train_data$genre <- ifelse(is.na(train_data$genre),"No_Genre", train_data$genre)

#train_data <- na.omit(train_data)
(cols <- unique(unlist(strsplit(train_data$genre, ","))))
cols <- gsub("[[:punct:]]", "", cols)
(cols <- str_trim(cols,side = "both"))
for (i in cols) train_data[[i]] <- grepl(i, train_data$genre)

train_data_1 <- train_data[,-c(1:19)]
train_data_1 <- ifelse(train_data_1 == "TRUE", 1, 0)

# selecting cols sorting to get top 6 most frequnt genres
(col_sums_train <- apply(X = train_data_1,2,sum))
sort(col_sums_train,decreasing = TRUE)

# binding the top 6 data columns
train_data_1 <- train_data_1[,c("pop","rock","mellow gold","soul","rap","soft rock","adult standards","country","dance pop","brill building pop","hip hop")]
train_data <- cbind(train_data[,c(1:19)],train_data_1)


# test data genre column splits to 1 and 0
test_data$genre<- gsub("\\[|\\]","",as.character(test_data$genre))
test_data$genre <- ifelse(test_data$genre == "","No_Genre", test_data$genre)
test_data$genre <- ifelse(is.na(test_data$genre),"No_Genre", test_data$genre)


(cols <- unique(unlist(strsplit(test_data$genre, ","))))
cols <- gsub("[[:punct:]]", "", cols)
(cols <- str_trim(cols,side = "both"))
for (i in cols) test_data[[i]] <- grepl(i, test_data$genre)

test_data_1 <- test_data[,-c(1:19)]
test_data_1 <- ifelse(test_data_1 == "TRUE", 1, 0)

# selecting cols sorting to get top 6 most frequnt genres
(col_sums_test <- apply(X = test_data_1,2,sum))
sort(col_sums_test,decreasing = TRUE)

# binding top 6 data columns
test_data_1 <- test_data_1[,c("pop","rock","mellow gold","soul","rap","soft rock","adult standards","country","dance pop","brill building pop","hip hop")]
test_data <- cbind(test_data[,c(1:18)],test_data_1)


#



# final data set to work with
skimr::skim(train_data)

train_data <- train_data[,-c(4,3)]
test_data <- test_data[,-c(4,3)]

train_data$track_explicit <- as.numeric(as.factor(train_data$track_explicit))
test_data$track_explicit <- as.numeric(as.factor(test_data$track_explicit))

names(train_data)[names(train_data) == "mellow gold"] <- "mellow_gold"
names(train_data)[names(train_data) == "soft rock"] <- "soft_rock"
names(train_data)[names(train_data) == "adult standards"] <- "adult_standards"
names(train_data)[names(train_data) == "dance pop"] <- "dance_pop"
names(train_data)[names(train_data) == "brill building pop"] <- "brill_building_pop"
names(train_data)[names(train_data) == "hip hop"] <- "hip_hop"


names(test_data)[names(test_data) == "mellow gold"] <- "mellow_gold"
names(test_data)[names(test_data) == "soft rock"] <- "soft_rock"
names(test_data)[names(test_data) == "adult standards"] <- "adult_standards"
names(test_data)[names(test_data) == "dance pop"] <- "dance_pop"
names(test_data)[names(test_data) == "brill building pop"] <- "brill_building_pop"
names(test_data)[names(test_data) == "hip hop"] <- "hip_hop"



```

# split data

```{r}
split <- createDataPartition(y = train_data$rating,p = 0.5,list = F,groups = 100)
train_s <- train_data[split,]
test_s <- train_data[-split,]

# str(train_s)
# train_s$genre<- gsub("\\[|\\]","",as.character(train_s$genre))
# train_s$genre <- ifelse(train_s$genre == "",NA, train_s$genre)
# train_s <- na.omit(train_s)
# (cols <- unique(unlist(strsplit(train_s$genre, ","))))
# cols <- gsub("[[:punct:]]", "", cols)
# (cols <- str_trim(cols,side = "both"))
# for (i in cols) train_s[[i]] <- grepl(i, train_s$genre)
# 
# train_s_1 <- train_s[,-c(1:19)]
# train_s_1 <- ifelse(train_s_1 == "TRUE", 1, 0)
# 
# train_s <- cbind(train_s[,c(1:19)],train_s_1)
# 
# 
# 
# test_s$genre<- gsub("\\[|\\]","",as.character(test_s$genre))
# test_s$genre <- ifelse(test_s$genre == "",NA, test_s$genre)
# test_s <- na.omit(test_s)
# (cols <- unique(unlist(strsplit(test_s$genre, ","))))
# cols <- gsub("[[:punct:]]", "", cols)
# (cols <- str_trim(cols,side = "both"))
# for (i in cols) test_s[[i]] <- grepl(i, test_s$genre)
# 
# test_s_1 <- test_s[,-c(1:19)]
# test_s_1 <- ifelse(test_s_1 == "TRUE", 1, 0)
# 
# test_s <- cbind(test_s[,c(1:19)],test_s_1)
```

# visualisation, data exploration

```{r}
gather(data = train_data[c(3,5:17)],key = "song_info",value = "info", -rating) |>
  ggplot(aes(x = info, y = rating)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~song_info,scales = "free")


ggcorrplot(cor(train_data[c(3,5:28)]) , type = "lower" , show.diag = T , colors = c("red","white","green") , lab = T, lab_size = 1)
```



# regessions

**taking all the variables except ID and performer as wrangling it is not making practical sense therefore would only get predictive power which would hug this model and not be generalisable to the world.**


```{r}

# model1 running a regression on data 
model1 = lm(rating ~ track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop, data=train_s)
summary(model1)
# train
pred1 <- predict(model1)
rmse1 <- sqrt(mean((pred1-train_s$rating)^2))
# test
pred1_test <- predict(model1,newdata=test_s)
rmse1_test <- sqrt(mean((pred1_test-test_s$rating)^2))


# Subset selection will run multiple models which will overheat the laptop


# forward selection
start_mod = lm(rating~1,data=train_s)
empty_mod = lm(rating~1,data=train_s)
full_mod = lm(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s)
forwardStepwise = step(start_mod,
                       scope=list(upper=full_mod,lower=empty_mod),
                       direction='forward')
forwardStepwise

# running a forward selection regression model
model2 <- lm(formula = rating ~ dance_pop + acousticness + rap + rock + 
    track_duration + loudness + energy + danceability + valence + 
    mellow_gold + brill_building_pop + pop + time_signature + 
    hip_hop + track_explicit + instrumentalness + tempo + liveness + 
    soul + adult_standards, data = train_s)
summary(model2)
# train
pred2 <- predict(model2)
rmse2 <- sqrt(mean((pred2-train_s$rating)^2))
# test
pred2_test <- predict(model2,newdata=test_s)
rmse2_test <- sqrt(mean((pred2_test-test_s$rating)^2))

# backword selection
start_mod = lm(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s)
empty_mod = lm(rating~1,data=train_s)
full_mod = lm(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s)
backwardStepwise = step(start_mod,
                        scope=list(upper=full_mod,lower=empty_mod),
                        direction='backward')
backwardStepwise

# running backward selection regression
model3 <- lm(formula = rating ~ track_duration + danceability + energy + 
    track_explicit + loudness + acousticness + instrumentalness + 
    liveness + valence + tempo + time_signature + pop + rock + 
    mellow_gold + soul + rap + adult_standards + dance_pop + 
    brill_building_pop + hip_hop, data = train_s)
summary(model3)
# train
pred3 <- predict(model3)
rmse3 <- sqrt(mean((pred3-train_s$rating)^2))
# test
pred3_test <- predict(model3,newdata=test_s)
rmse3_test <- sqrt(mean((pred3_test-test_s$rating)^2))


# Stepwise variable selection

start_mod = lm(rating~1,data=train_s)
empty_mod = lm(rating~1,data=train_s)
full_mod = lm(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s)
stepwise = step(start_mod,
                       scope=list(upper=full_mod,lower=empty_mod),
                       direction='both')
stepwise

model4 <- lm(formula = rating ~ dance_pop + acousticness + rap + rock + 
    track_duration + loudness + energy + danceability + valence + 
    mellow_gold + brill_building_pop + pop + time_signature + 
    hip_hop + track_explicit + instrumentalness + tempo + liveness + 
    soul + adult_standards, data = train_s)
summary(model4)
# train
pred4 <- predict(model4)
rmse4 <- sqrt(mean((pred4-train_s$rating)^2))
# test
pred4_test <- predict(model4,newdata=test_s)
rmse4_test <- sqrt(mean((pred4_test-test_s$rating)^2))

library(mgcv)
model5 <- gam(formula = rating ~ dance_pop + acousticness + rap + rock + 
    s(track_duration) + loudness + energy + danceability + valence + 
    mellow_gold + brill_building_pop + pop + time_signature + 
    hip_hop + track_explicit + instrumentalness + s(tempo) + liveness + 
    soul + adult_standards, data = train_s,method = "REML") 
summary(model5)
# train
pred5 <- predict(model5)
rmse5 <- sqrt(mean((pred5-train_s$rating)^2))
# test
pred5_test <- predict(model5,newdata=test_s)
rmse5_test <- sqrt(mean((pred5_test-test_s$rating)^2))


# run a lasso regression 
library(glmnet)

x = model.matrix(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,train_s)
y = train_s$rating
set.seed(15)
lasso <- cv.glmnet(x=x,y = y,alpha = 1) # lasso when alpha is 1
coef(lasso, s = lasso$lambda.1se) %>%
  round(4)

# lasso shrinked regression

model6 <- lm(rating~danceability+energy+track_explicit+loudness+acousticness+instrumentalness+liveness+valence+time_signature+pop+rock+mellow_gold+soul+rap+dance_pop+brill_building_pop,data=train_s)
summary(model6)
# train
pred6 <- predict(model6)
rmse6 <- sqrt(mean((pred6-train_s$rating)^2))
# test
pred6_test <- predict(model6,newdata=test_s)
rmse6_test <- sqrt(mean((pred6_test-test_s$rating)^2))


summary_reg <- data.frame(Name = 1:6,rmse_train = c(rmse1,rmse2,rmse3,rmse4,rmse5,rmse6), rmse_test = c(rmse1_test,rmse2_test,rmse3_test,rmse4_test,rmse5_test,rmse6_test) )
summary_reg


gather(data=summary_reg,key = "rmse", value= "score", -Name) |> ggplot(aes(x=Name,y=score)) + geom_point() + facet_wrap(~rmse,ncol=1)

```


# Trees

```{r}
library(rpart); library(rpart.plot)

# regular trees without tuning
model7 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova')
summary(model7)
# train
pred7 <- predict(model7)
rmse7 <- sqrt(mean((pred7-train_s$rating)^2))
# test
pred7_test <- predict(model7,newdata=test_s)
rmse7_test <- sqrt(mean((pred7_test-test_s$rating)^2))

# trying out some parameters
model8 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova',control = rpart.control(minsplit = 100))
summary(model8)
# train
pred8 <- predict(model8)
rmse8 <- sqrt(mean((pred8-train_s$rating)^2))
# test
pred8_test <- predict(model8,newdata=test_s)
rmse8_test <- sqrt(mean((pred8_test-test_s$rating)^2))

# another one
model9 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova',control = rpart.control(minbucket = 10))
summary(model9)
# train
pred9 <- predict(model9)
rmse9 <- sqrt(mean((pred9-train_s$rating)^2))
# test
pred9_test <- predict(model9,newdata=test_s)
rmse9_test <- sqrt(mean((pred9_test-test_s$rating)^2))

# another one
model10 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova',control = rpart.control(maxdepth = 30))
summary(model10)
# train
pred10 <- predict(model10)
rmse10 <- sqrt(mean((pred10-train_s$rating)^2))
# test
pred10_test <- predict(model10,newdata=test_s)
rmse10_test <- sqrt(mean((pred10_test-test_s$rating)^2))


# another one
model11 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova',control = rpart.control(cp = 0.00001))
summary(model11)
# train
pred11 <- predict(model11)
rmse11 <- sqrt(mean((pred11-train_s$rating)^2))
# test
pred11_test <- predict(model11,newdata=test_s)
rmse11_test <- sqrt(mean((pred11_test-test_s$rating)^2))

# mixing all the tunes

model12 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data = train_s, method = 'anova',control = rpart.control(cp = 0.00001,minsplit = 100,minbucket = 10,maxdepth = 30))
summary(model12)
# train
pred12 <- predict(model12)
rmse12 <- sqrt(mean((pred12-train_s$rating)^2))
# test
pred12_test <- predict(model12,newdata=test_s)
rmse12_test <- sqrt(mean((pred12_test-test_s$rating)^2))


```

## TUNING TREES

```{r}
# tuning decision trees
trControl = trainControl(method='cv',number = 5)
tuneGrid = expand.grid(cp = seq(from = 0.001,to = 0.1,by = 0.001))
set.seed(617)
cvModel13 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
                data=train_s,
                method="rpart",
                trControl = trControl,
                tuneGrid = tuneGrid)

model13 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,cp = cvModel13$bestTune$cp)
summary(model13)
# train
pred13 <- predict(model13)
rmse13 <- sqrt(mean((pred13-train_s$rating)^2))
# test
pred13_test <- predict(model13,newdata=test_s)
rmse13_test <- sqrt(mean((pred13_test-test_s$rating)^2))



# tuning multiple parameters for decision trees
# 
# trControl = trainControl(method='cv',number = 5)
# tuneGrid = expand.grid(cp = seq(from = 0.001,to = 0.1,by = 0.001),minsplit = c(10,20,30,50,100),minbucket = c(20,10,5),maxdepth = c(10,15,30))
# set.seed(617)
# cvModel14 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
#                 data=train_s,
#                 method="rpart",
#                 trControl = trControl,
#                 tuneGrid = tuneGrid)
# 
# model13 = rpart(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,cp = cvModel13$bestTune$cp)
# summary(model13)
# # train
# pred13 <- predict(model13)
# rmse13 <- sqrt(mean((pred13-train_s$rating)^2))
# # test
# pred13_test <- predict(model13,newdata=test_s)
# rmse13_test <- sqrt(mean((pred13_test-test_s$rating)^2))


# bagging 
library(ipred)
set.seed(15) 
model14 = bagging(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
              data = train_s, 
              nbagg = 1000)

summary(model14)
# train
pred14 <- predict(model14)
rmse14 <- sqrt(mean((pred14-train_s$rating)^2))
# test
pred14_test <- predict(model14,newdata=test_s)
rmse14_test <- sqrt(mean((pred14_test-test_s$rating)^2))


# bagging using random forests
library(randomForest)
set.seed(15)
model15 = randomForest(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
              data = train_s, 
              mtry= ncol(train_s)-1,ntree=1000)
summary(model15)
# train
pred15 <- predict(model15)
rmse15 <- sqrt(mean((pred15-train_s$rating)^2))
# test
pred15_test <- predict(model15,newdata=test_s)
rmse15_test <- sqrt(mean((pred15_test-test_s$rating)^2))

# random forest model
set.seed(15)
model16 = randomForest(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
              data = train_s, 
              ,ntree=1000)
summary(model16)
# train
pred16 <- predict(model16)
rmse16 <- sqrt(mean((pred16-train_s$rating)^2))
# test
pred16_test <- predict(model16,newdata=test_s)
rmse16_test <- sqrt(mean((pred16_test-test_s$rating)^2))

# tuned Random forest model

trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:10)
set.seed(15)
cvModel17 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,
                 method="rf",ntree=500,trControl=trControl,tuneGrid=tuneGrid )

model17 = randomForest(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,ntree = 500,mtry=cvModel17$bestTune$mtry)
summary(model17)
# train
pred17 <- predict(model17)
rmse17 <- sqrt(mean((pred17-train_s$rating)^2))
# test
pred17_test <- predict(model17,newdata=test_s)
rmse17_test <- sqrt(mean((pred17_test-test_s$rating)^2))

# forest with ranger
library(ranger)
model18 = ranger(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,num.trees = 1000)
summary(model18)
# train
pred18 <- predict(model18,data=train_s,num.trees=1000)
rmse18 <- sqrt(mean((pred18$predictions-train_s$rating)^2))
# test
pred18_test <- predict(model18,data=test_s,num.trees=1000)
rmse18_test <- sqrt(mean((pred18_test$predictions-test_s$rating)^2))


# tuned forest ranger
trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:4, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(2,5,10,15,20,25))
cvModel19 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
                data=train_s,
                method="ranger",
                num.trees=200,
                trControl=trControl,
                tuneGrid=tuneGrid )


set.seed(15)
model19 = ranger(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,num.trees = 200,mtry=cvModel19$bestTune$mtry, 
                          min.node.size = cvModel19$bestTune$min.node.size, 
                          splitrule = cvModel19$bestTune$splitrule)
summary(model19)
# train
pred19 <- predict(model19,data=train_s,num.trees=200)
rmse19 <- sqrt(mean((pred19$predictions-train_s$rating)^2))
# test
pred19_test <- predict(model19,data=test_s,num.trees=200)
rmse19_test <- sqrt(mean((pred19_test$predictions-test_s$rating)^2))


# boosting models
# gbm
library(gbm)
set.seed(15)
model20 = gbm(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
            data=train_s,
            distribution="gaussian",
            n.trees = 200,
            interaction.depth = 2,
            shrinkage = 0.01)


summary(model20)
# train
pred20 <- predict(model20,n.trees=200)
rmse20 <- sqrt(mean((pred20-train_s$rating)^2))
# test
pred20_test <- predict(model20,newdata=test_s,n.trees=200)
rmse20_test <- sqrt(mean((pred20_test-test_s$rating)^2))


# xgboost
library(vtreat)
trt = designTreatmentsZ(dframe = train_s,
                        varlist = names(train_s)[c(3,5:16,18:28)])
newvars = trt$scoreFrame[trt$scoreFrame$code%in% c('clean','lev'),'varName']

train_input = prepare(treatmentplan = trt, 
                      dframe = train_s,
                      varRestriction = newvars)
test_input = prepare(treatmentplan = trt, 
                     dframe = test_s,
                     varRestriction = newvars)
main_tes = prepare(treatmentplan = trt,dframe=test_data)
head(train_input)

library(xgboost); library(caret)
set.seed(15)
tune_nrounds21 = xgb.cv(data=as.matrix(train_input), 
                      label = train_s$rating,
                      nrounds=250,
                      nfold = 5,
                      verbose = 0)

which.min(tune_nrounds$evaluation_log$test_rmse_mean)

model21= xgboost(data=as.matrix(train_input), 
                  label = train_s$rating,
                  nrounds=11, # after tuned rounds
                  verbose = 0)
summary(model21)
# train
pred21 <- predict(model21,newdata = as.matrix(train_input))
rmse21 <- sqrt(mean((pred21-train_s$rating)^2))
# test
pred21_test <- predict(model21,newdata=as.matrix(test_input))
rmse21_test <- sqrt(mean((pred21_test-test_s$rating)^2))


# tuned forest ranger with incresed trees
trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:4, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(2,5,10,15,20,25))
cvModel22 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
                data=train_s,
                method="ranger",
                num.trees=2000,
                trControl=trControl,
                tuneGrid=tuneGrid )


set.seed(15)
model22 = ranger(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,num.trees = 2000,mtry=cvModel22$bestTune$mtry, 
                          min.node.size = cvModel22$bestTune$min.node.size, 
                          splitrule = cvModel22$bestTune$splitrule)
summary(model22)
# train
pred22 <- predict(model22,data=train_s,num.trees=2000)
rmse22 <- sqrt(mean((pred22$predictions-train_s$rating)^2))
# test
pred22_test <- predict(model22,data=test_s,num.trees=2000)
rmse22_test <- sqrt(mean((pred22_test$predictions-test_s$rating)^2))


# tuned ranger with more parameters

trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry=1:10, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(1,2,5,10,15,20,25))
cvModel23 = train(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,
                data=train_s,
                method="ranger",
                num.trees=2000,
                trControl=trControl,
                tuneGrid=tuneGrid )


set.seed(15)
model23 = ranger(rating~track_duration+ danceability+ energy+ key+ mode + track_explicit + loudness+speechiness+acousticness+instrumentalness+liveness+valence+tempo+time_signature+pop+rock+mellow_gold+soul+rap+soft_rock+adult_standards+country+dance_pop+brill_building_pop+hip_hop,data=train_s,num.trees = 2000,mtry=cvModel23$bestTune$mtry, 
                          min.node.size = cvModel23$bestTune$min.node.size, 
                          splitrule = cvModel23$bestTune$splitrule)
summary(model23)
# train
pred23 <- predict(model23,data=train_s,num.trees=2000)
rmse23 <- sqrt(mean((pred23$predictions-train_s$rating)^2))
# test
pred23_test <- predict(model23,data=test_s,num.trees=2000)
rmse23_test <- sqrt(mean((pred23_test$predictions-test_s$rating)^2))

```

# visualise models

how all the models performed

```{r}
summary_reg <- data.frame(Name = c("Regression","forward_model","Backward_model","stepwise_model","gam","lasso","Decision_tree","tree_1","tree_2","tree_3","tree_4","tree_para_mix","tuned_tree","Bagging_ipred","Bagging_rf","rf","tuned_rf","rf_ranger","tuned_rf_ranger","boosting_gbm","Xgboost","Tuned_rf_pro","tuned_rf_pro_max"),rmse_train = c(rmse1,rmse2,rmse3,rmse4,rmse5,rmse6,rmse7,rmse8,rmse9,rmse10,rmse11,rmse12,rmse13,rmse14,rmse15,rmse16,rmse17,rmse18,rmse19,rmse20,rmse21,rmse22,rmse23), rmse_test = c(rmse1_test,rmse2_test,rmse3_test,rmse4_test,rmse5_test,rmse6_test,rmse7_test,rmse8_test,rmse9_test,rmse10_test,rmse11_test,rmse12_test,rmse13_test,rmse14_test,rmse15_test,rmse16_test,rmse17_test,rmse18_test,rmse19_test,rmse20_test,rmse21_test,rmse22_test,rmse23_test) )
summary_reg


gather(data=summary_reg,key = "rmse", value= "score", -Name) |> ggplot(aes(x=Name,y=score)) + geom_point() + facet_wrap(~rmse,ncol=1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Main test data predictions

all the models predictions with the main test data
```{r}
pred1_test_main <- predict(model1,newdata = test_data)
pred2_test_main <- predict(model2,newdata = test_data)
pred3_test_main <- predict(model3,newdata = test_data)
pred4_test_main <- predict(model4,newdata = test_data)
pred5_test_main <- predict(model5,newdata = test_data)
pred6_test_main <- predict(model6,newdata = test_data)
pred7_test_main <- predict(model7,newdata = test_data)
pred8_test_main <- predict(model8,newdata = test_data)
pred9_test_main <- predict(model9,newdata = test_data)
pred10_test_main <- predict(model10,newdata = test_data)
pred11_test_main <- predict(model11,newdata = test_data)
pred12_test_main <- predict(model12,newdata = test_data)
pred13_test_main <- predict(model13,newdata = test_data)
pred14_test_main <- predict(model14,newdata = test_data)
pred15_test_main <- predict(model15,newdata = test_data)
pred16_test_main <- predict(model16,newdata = test_data)
pred17_test_main <- predict(model17,newdata = test_data)
pred18_test_main <- predict(model18,data = test_data,num.trees=1000)
pred19_test_main <- predict(model19,data = test_data,num.trees=200)
pred20_test_main <- predict(model20,newdata = test_data)
pred21_test_main <- predict(model21,newdata = as.matrix(main_tes))
pred22_test_main <- predict(model22,data = test_data,num.trees=2000)
pred23_test_main <- predict(model23,data = test_data,num.trees = 2000)





```



# submissions

better to comment out this section as it will only run in my local laptop

```{r}
# Submission 1
submissionFile = data.frame(id = test_data$id, rating = pred1_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model1.csv',row.names = F)

# Submission 2
submissionFile = data.frame(id = test_data$id, rating = pred2_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model2.csv',row.names = F)

# Submission 3
submissionFile = data.frame(id = test_data$id, rating = pred3_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model3.csv',row.names = F)

# Submission 4
submissionFile = data.frame(id = test_data$id, rating = pred4_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model4.csv',row.names = F)

# Submission 5
submissionFile = data.frame(id = test_data$id, rating = pred5_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model5.csv',row.names = F)

# Submission 6
submissionFile = data.frame(id = test_data$id, rating = pred6_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model6.csv',row.names = F)

# Submission 7
submissionFile = data.frame(id = test_data$id, rating = pred7_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model7.csv',row.names = F)

# Submission 8
submissionFile = data.frame(id = test_data$id, rating = pred8_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model8.csv',row.names = F)

# Submission 9
submissionFile = data.frame(id = test_data$id, rating = pred9_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model9.csv',row.names = F)

# Submission 10
submissionFile = data.frame(id = test_data$id, rating = pred10_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model10.csv',row.names = F)

# Submission 11
submissionFile = data.frame(id = test_data$id, rating = pred11_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model11.csv',row.names = F)

# Submission 12
submissionFile = data.frame(id = test_data$id, rating = pred12_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model12.csv',row.names = F)

# Submission 13
submissionFile = data.frame(id = test_data$id, rating = pred13_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model13.csv',row.names = F)

# Submission 14
submissionFile = data.frame(id = test_data$id, rating = pred14_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model14.csv',row.names = F)

# Submission 15
submissionFile = data.frame(id = test_data$id, rating = pred15_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model15.csv',row.names = F)

# Submission 16
submissionFile = data.frame(id = test_data$id, rating = pred16_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model16.csv',row.names = F)

# Submission 17
submissionFile = data.frame(id = test_data$id, rating = pred17_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model17.csv',row.names = F)

# Submission 18
submissionFile = data.frame(id = test_data$id, rating = pred18_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model18.csv',row.names = F)

# Submission 19
submissionFile = data.frame(id = test_data$id, rating = pred19_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model19.csv',row.names = F)

# Submission 20
submissionFile = data.frame(id = test_data$id, rating = pred20_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model20.csv',row.names = F)

# Submission 21
submissionFile = data.frame(id = test_data$id, rating = pred21_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model21.csv',row.names = F)

# Submission 22
submissionFile = data.frame(id = test_data$id, rating = pred22_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model22.csv',row.names = F)

# Submission 23
submissionFile = data.frame(id = test_data$id, rating = pred23_test_main)
write.csv(submissionFile, '/Users/ishujaswani/downloads/model23.csv',row.names = F)

```

